---
title: "Findings from the ACFR Dataset"
output: 
  html_document: 
    toc: true
    toc_float: true
    number_sections: true

author: "Mariana F Trujillo"
date: "April 2024"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(rio)
library(ggridges)
library(ggpattern)
library(ggdist)
library(ggbeeswarm)
library(ggstream)
library(ggpubr)
library(scales)
library(knitr)
library(reactable)
library(htmltools)
library(zoo)
library(quantmod)
library(PerformanceAnalytics)
library(quadprog)
library(janitor)
library(kableExtra)
library(patchwork)
library(showtext)
library(RColorBrewer)

States3y <- read.csv("all_states_3years.csv")
Cities3y <- read.csv("top100_cities_3years.csv")
Counties3y <- read.csv("top100_counties_3years.csv")
SD3y <- read.csv("top100_sd_3years.csv")

source("APR functions.R")
source("Reason theme.R")

knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  fig.align = "center",
  cache = FALSE
)
```

```{r, echo=FALSE}
#Add per capita
States3y <- States3y %>%
  mutate(npl_per_capita = net_pension_liability / population) 

Cities3y  <- Cities3y %>%
  mutate(npl_per_capita = net_pension_liability / population) 

Counties3y <- Counties3y %>%
  mutate(npl_per_capita = net_pension_liability / population) 

SD3y <- SD3y %>%
  mutate(npl_per_student= net_pension_liability / students) 

```

```{r, echo=FALSE}

Combined3y <- bind_rows(
  mutate(States3y, dataset = "States"),
  mutate(Cities3y, dataset = "Cities"),
  mutate(Counties3y, dataset = "Counties"),
  mutate(SD3y, dataset = "School Districts")
)

total_unfunded_liabilities <- Combined3y %>%
  group_by(year, dataset) %>%
  summarise(total_unfunded_liabilities = sum(net_pension_liability, na.rm = TRUE), .groups = "drop")

total_unfunded_liabilities$dataset <- factor(total_unfunded_liabilities$dataset, levels = c("States", "Cities", "Counties", "School Districts"))

options(repr.plot.width = 10, repr.plot.height = 6)

```

``` {r, echo=FALSE}
States_Total_NPL <- States3y %>%
  group_by(year) %>%
  summarise(net_opeb_liability = sum(net_opeb_liability, na.rm = TRUE),
            net_pension_liability = sum(net_pension_liability, na.rm = TRUE))

States_Total_NPL_change <- States_Total_NPL %>%
  mutate(rate_of_change = (net_opeb_liability - lag(net_opeb_liability)) / lag(net_opeb_liability) * 100)


SD_Total_NPL <- SD3y %>%
  group_by(year) %>%
  summarise(net_opeb_liability = sum(net_opeb_liability, na.rm = TRUE),
            net_pension_liability = sum(net_pension_liability, na.rm = TRUE))

SD_Total_NPL_change <- SD_Total_NPL %>%
  mutate(rate_of_change_opeb = (net_opeb_liability - lag(net_opeb_liability)) / lag(net_opeb_liability) * 100)


Cities_Total_NPL <- Cities3y %>%
  group_by(year) %>%
  summarise(net_opeb_liability = sum(net_opeb_liability, na.rm = TRUE),
            net_pension_liability = sum(net_pension_liability, na.rm = TRUE))

Cities_Total_NPLchange <- Cities_Total_NPL %>%
  mutate(rate_of_change_opeb = (net_opeb_liability - lag(net_opeb_liability)) / lag(net_opeb_liability) * 100)


Counties_Total_NPL <- Counties3y %>%
  group_by(year) %>%
  summarise(net_opeb_liability = sum(net_opeb_liability, na.rm = TRUE),
            net_pension_liability = sum(net_pension_liability, na.rm = TRUE))

Counties_Total_NPLchange <- Counties_Total_NPL %>%
  mutate(rate_of_change_opeb = (net_opeb_liability - lag(net_opeb_liability)) / lag(net_opeb_liability) * 100)

combined_change_data <- bind_rows(
  States_Total_NPL_change %>% mutate(Location = "State"),
  SD_Total_NPL_change %>% mutate(Location = "SD"),
  Cities_Total_NPLchange %>% mutate(Location = "City"),
  Counties_Total_NPLchange %>% mutate(Location = "County")
) %>%
select(rate_of_change_opeb, Location, year, net_opeb_liability, net_pension_liability)

orange_colors <- brewer.pal(3, "YlOrBr")


```

```{r, echo=FALSE}
States3y <- States3y %>%
  mutate(npl_per_capita = net_pension_liability / population) 

Cities3y  <- Cities3y %>%
  mutate(npl_per_capita = net_pension_liability / population) 

Counties3y <- Counties3y %>%
  mutate(npl_per_capita = net_pension_liability / population) 

SD3y <- SD3y %>%
  mutate(npl_per_student= net_pension_liability / students) 

```

# Trends

## Unfunded Liabilities 


```{r, echo=FALSE}

States3y_sorted <- States3y %>%
  arrange(name, year)

States_NPL_20to22diff <- States3y_sorted %>%
  group_by(name) %>%
  mutate(
    OPEB_nominal_difference_NPL20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_NPL20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100,
    OPEB_nominal_difference_L20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_L20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100 )

Cities3y_sorted <- Cities3y %>%
  arrange(name, year)

Cities_NPL_20to22diff <- Cities3y_sorted %>%
  group_by(name) %>%
  mutate(
    OPEB_nominal_difference_NPL20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_NPL20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100,
    OPEB_nominal_difference_L20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_L20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100 )

Counties3y_sorted <- Counties3y %>%
  arrange(name, year)

Counties_NPL_20to22diff <- Counties3y_sorted %>%
  group_by(name) %>%
  mutate(
    OPEB_nominal_difference_NPL20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_NPL20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100,
    OPEB_nominal_difference_L20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_L20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100 )

SD3y_sorted <- SD3y %>%
  arrange(name, year)

SD_NPL_20to22diff <- SD3y_sorted %>%
  group_by(name) %>%
  mutate(
    OPEB_nominal_difference_NPL20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_NPL20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100,
    OPEB_nominal_difference_L20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_L20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100 )


```


```{r, echo=FALSE}

total_OPEB_liabilities <- Combined3y %>%
  group_by(dataset, year) %>%
  summarise(total_OPEB_liabilities = sum(net_opeb_liability, na.rm = TRUE), .groups = "drop")

total_OPEB_liabilities <- total_OPEB_liabilities %>%
  group_by(dataset) %>%
  summarise(
    first_total_OPEB = first(total_OPEB_liabilities),
    last_total_OPEB = last(total_OPEB_liabilities),
    OPEB_nominal_difference_NPL20to22 = last_total_OPEB - first_total_OPEB,
    OPEB_percent_difference_NPL20to22 = ((last_total_OPEB - first_total_OPEB) / first_total_OPEB) * 100,
    OPEB_nominal_difference_L20to22 = last_total_OPEB - first_total_OPEB,
    OPEB_percent_difference_L20to22 = ((last_total_OPEB - first_total_OPEB) / first_total_OPEB) * 100,
    .groups = "drop")


```

```{r, echo=FALSE}

Combined3y <- bind_rows(
  mutate(States3y, dataset = "States"),
  mutate(Cities3y, dataset = "Cities"),
  mutate(Counties3y, dataset = "Counties"),
  mutate(SD3y, dataset = "School Districts")
)

total_OPEB_liabilities <- Combined3y %>%
  group_by(dataset, year) %>%
  summarise(total_OPEB_liabilities = sum(net_opeb_liability, na.rm = TRUE), .groups = "drop")

total_OPEB_liabilities <- total_OPEB_liabilities %>%
  group_by(dataset) %>%
  summarise(
    first_total_OPEB = first(total_OPEB_liabilities),
    last_total_OPEB = last(total_OPEB_liabilities),
    OPEB_nominal_difference_NPL20to22 = last_total_OPEB - first_total_OPEB,
    OPEB_percent_difference_NPL20to22 = ((last_total_OPEB - first_total_OPEB) / first_total_OPEB) * 100,
    OPEB_nominal_difference_L20to22 = last_total_OPEB - first_total_OPEB,
    OPEB_percent_difference_L20to22 = ((last_total_OPEB - first_total_OPEB) / first_total_OPEB) * 100,
    .groups = "drop")


```


```{r, echo=FALSE}
combined_change_data %>%
  ggplot(aes(x = Location, y = net_opeb_liability, fill = factor(year))) +
  geom_col(position = "dodge") +
  geom_text(aes(label = scales::dollar(net_opeb_liability)), 
            position = position_dodge(width = 0.9),
            vjust = -0.5, 
            size = 3, 
            show.legend = FALSE,
            hjust = -0.2) +  # Adjust position to sit outside the graph
  labs(
    title = "Change in Unfunded OPEB Liabilities by Entity",
    x = "Entity",
    y = "Net OPEB Liability",
    fill = "Year"
  ) +
  scale_fill_manual(values = orange_colors) +
  theme_minimal() +
  scale_y_continuous(labels = scales::dollar_format(prefix = "$", scale = 1, accuracy = 2)) +  
  coord_flip() +
  theme_reason ()

```

```{r, echo=FALSE}
combined_change_data %>%
  ggplot(aes(x = Location, y = net_pension_liability, fill = factor(year))) +
  geom_col(position = "dodge") +
  geom_text(aes(label = scales::dollar(net_opeb_liability)), 
            position = position_dodge(width = 0.9),
            vjust = -0.5, 
            size = 3, 
            show.legend = FALSE,
            hjust = -0.2) +  # Adjust position to sit outside the graph
  labs(
    title = "Change in Unfunded Pension Liabilities by Entity",
    x = "Entity",
    y = "Net OPEB Liability",
    fill = "Year"
  ) +
  scale_fill_manual(values = orange_colors) +
  theme_minimal() +
  scale_y_continuous(labels = scales::dollar_format(prefix = "$", scale = 1, accuracy = 2)) +  
  coord_flip() +
  theme_reason ()

```



## OPEB Liabilities 


### How West Virginia and Alaska abolished their Unfunded OPEB liability 


``` {r, echo=FALSE}
selected_columns <- c("name", "net_pension_liability", "npl_per_capita", "net_opeb_liability",  "revenues")

subset_states3y <- States3y %>%
  filter(year == 2022) %>%
  filter(year == 2022 & name %in% c("west virginia", "alaska")) %>%
  select(all_of(selected_columns))

reactable(
  data = subset_states3y,
  filterable = FALSE,
  defaultSorted = list(net_pension_liability = "desc"),
  columns = list(
    name = colDef(name = "State"),
    net_pension_liability = colDef(name = "Net Pension Liability", cell = function(value) sprintf("$ %s", format(value, big.mark = ","))),
    npl_per_capita = colDef(name = "Net Pension Liability Per Capita", cell = function(value) sprintf("$ %s", format(value, big.mark = ","))),
    net_opeb_liability = colDef(name = "Net OPEB Liability", cell = function(value) sprintf("$ %s", format(value, big.mark = ","))),
    revenues = colDef(name = "Revenues", cell = function(value) sprintf("$ %s", format(value, big.mark = ",")))
  ),
  defaultPageSize = 10,
  searchable = FALSE
)

```

West Virginia: Lump sump legislature comtributions 
-The state reserved  enough money aside and slowly paid off their net OPEB liabilities.  
-Senate Bill No. 419 mandating contributions
-Net OPEB asset of $23,841,000

Alaska: Pension ADEC
-Alaska has actuarially determined OPEB employee contribution rate.
-A set statutory employer contribution rate (Alaska, 12%), and any difference between the actuarially determined contribution rate and the statutory employer effective rate is paid by the State. In 2022 the difference was of 1% of payroll.
-Net OPEB assets of $1B in 2023 !


### The general fall in OPEB liabilities 



```{r, echo=FALSE}

States3y_sorted <- States3y %>%
  arrange(name, year)

States_NPL_20to22diff <- States3y_sorted %>%
  group_by(name) %>%
  mutate(
    OPEB_nominal_difference_NPL20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_NPL20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100,
    OPEB_nominal_difference_L20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_L20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100 )

Cities3y_sorted <- Cities3y %>%
  arrange(name, year)

Cities_NPL_20to22diff <- Cities3y_sorted %>%
  group_by(name) %>%
  mutate(
    OPEB_nominal_difference_NPL20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_NPL20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100,
    OPEB_nominal_difference_L20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_L20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100 )

Counties3y_sorted <- Counties3y %>%
  arrange(name, year)

Counties_NPL_20to22diff <- Counties3y_sorted %>%
  group_by(name) %>%
  mutate(
    OPEB_nominal_difference_NPL20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_NPL20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100,
    OPEB_nominal_difference_L20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_L20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100 )

SD3y_sorted <- SD3y %>%
  arrange(name, year)

SD_NPL_20to22diff <- SD3y_sorted %>%
  group_by(name) %>%
  mutate(
    OPEB_nominal_difference_NPL20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_NPL20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100,
    OPEB_nominal_difference_L20to22 = last(net_opeb_liability) - first(net_opeb_liability),
    OPEB_percent_difference_L20to22 = (last(net_opeb_liability) - first(net_opeb_liability)) / first(net_opeb_liability) * 100 )


```



```{r, echo=FALSE}
combined_change_data %>%
  ggplot(aes(x = Location, y = net_opeb_liability, fill = factor(year))) +
  geom_col(position = "dodge") +
  geom_text(aes(label = scales::dollar(net_opeb_liability)), 
            position = position_dodge(width = 0.9),
            vjust = -0.5, 
            size = 3, 
            show.legend = FALSE,
            hjust = -0.2) +  # Adjust position to sit outside the graph
  labs(
    title = "Change in Unfunded Liabilities by Entity",
    x = "Entity",
    y = "Net OPEB Liability",
    fill = "Year"
  ) +
  scale_fill_manual(values = orange_colors) +
  theme_minimal() +
  scale_y_continuous(labels = scales::dollar_format(prefix = "$", scale = 1, accuracy = 2)) +  
  coord_flip() +
  theme_reason ()

```

```{r, echo=FALSE}

selected_total_OPEB_liabilities <- total_OPEB_liabilities %>%
  select(dataset, OPEB_nominal_difference_L20to22, OPEB_percent_difference_L20to22, last_total_OPEB) %>%
  mutate(
    OPEB_nominal_difference_L20to22 = dollar(OPEB_nominal_difference_L20to22),
    OPEB_percent_difference_L20to22 = percent(OPEB_percent_difference_L20to22 / 100, accuracy = 0.01),
    last_total_OPEB = dollar(last_total_OPEB)
  )


kable(selected_total_OPEB_liabilities, 
      col.names = c("Entuty","$ Difference 20-22", "% Difference 20-22", "2022 Unfunded OPEB Liabilities"), 
      caption = "Change in Total OPEB Unfunded Liabilities by Entity", 
      format = "html") %>%  
  kable_styling(full_width = FALSE)

```


# Case Studies





